
# Data Management Toolbox


## Joins and Relates Toolset

## Projections and Transformations Toolset

Quite often in GIS, users found themselves before the necessity of transforming
(reprojecting) the coordinate system of a dataset (vector or raster). In fact this
is one of the most commonly used operations. In ArcGIS pro the aforementioned
action is performed using the following tools:

- `Project (Data Management)` -> Projects __spatial data__ from one coordinate system to another.
- `Project Raster (Data Management)` -> Transforms a __raster dataset__ from one coordinate system to another.

The question now is how we can perform similar operations with R. 

### Reproject vector data in R

For our example we will use the following dataset.

- **bezirke.gpkg** -> Dataset that depicts the Municipal structure in Switzerland

**R** and more specifically **sf package** offers options (functions) not only for 
transforming the coordinate system of a dataset, but also for identifying the
current existing one. Let's dive in :)

```{r echo=TRUE, message=TRUE, warning=TRUE}
# Importing the dataset
library(sf)
bezirke <- read_sf("sample_data/other/bezirke.gpkg") 
```

```{r echo=TRUE, message=TRUE, warning=TRUE}
# Visualising the imported dataset
library(ggplot2)
ggplot() + geom_sf(data = bezirke)
```

In order to identify the current coordinate system of a dataset, **sf package**
offers the function _st_crs_. The main usage of the function is to retrieve the
coordinate reference system from an sf object.

```{r echo=TRUE, message=TRUE, warning=TRUE}
# retrieving the coordinate of the imported layer
st_crs(bezirke)
```
As an output we received that there is no assigned coordinate system to this
specific dataset. _st_set_crs_ function helps us assigning a coordinate system
to a dataset.

```{r echo=TRUE, message=TRUE, warning=TRUE}
# Assigning the World Geodetic System (WGS84) as a coordinate system to our dataset
bezirke <- st_set_crs(bezirke, 4326)
```
In the function above, the number 4326 represents the __EPSG Geodetic Parameter Dataset (EPSG)__
of the respective coordinate system. **EPSG** is a public registry of geodetic
datums, spatial reference systems, Earth ellipsoids, coordinate transformations
and related units of measurement.  

```{r echo=TRUE, message=TRUE, warning=TRUE}
st_crs(bezirke)
```
After rerunning the _st_crs_ function, we are getting now that the assigned
coordinate system of the dataset is World Geodetic System (WGS84).  
As a next, we might want to transform the dataset to the Swiss coordinate system.
The respective **EPSG** for code for the latest Swiss coordinate system
(CH1903+LV95) is 2056. For this purpose we use the _st_transform_ function of the **sf package**.

```{r echo=TRUE, message=TRUE, warning=TRUE}
bezirke_swiss <- st_transform(bezirke, 2056)

# retrieve the coordinate system
st_crs(bezirke_swiss)
```


### Reproject raster data in R

















## Topology Toolset

## Sampling Toolset

### Generate Tessellation

- Hexagon
- Tranverse Hexagon
- Square
- Diamond
- Triangle


### Create Fishnet

### Create Random Points

### Generate Points Along Lines

### Generate Rectangles Along Lines

### Generate Transsects Along Lines

## Raster Toolset

<!-- Yes, these are toolsets within toolsets -->
### Mosaic Dataset Toolset

### Ortho Mapping Toolset

### Raster Dataset Toolset

### Raster Processing Toolset

#### Clip Raster

#### Composite Bands

#### Compute Pansharpen Weights

#### Create Ortho Corrected Raster Dataset

#### Create Pansharpened Raster Dataset

#### Extract Subdataset

#### Generate Table From Raster Function

#### Raster To DTED

#### Resample

#### Split Raster


### Raster Properties Toolset


