
# Data Management Toolbox


## Joins and Relates Toolset

## Projections and Transformations Toolset

Quite often in GIS, users found themselves before the necessity of transforming
(reprojecting) the coordinate system of a dataset (vector or raster). In fact this
is one of the most commonly used operations. In ArcGIS pro the aforementioned
action is performed using the following tools:

- `Project (Data Management)` -> Projects __spatial data__ from one coordinate system to another.
- `Project Raster (Data Management)` -> Transforms a __raster dataset__ from one coordinate system to another.

The question now is how we can perform similar operations with R. 

### Reproject vector data in R

For our example we will use the following data.

- **kantone.gpkg**: Dataset that depicts the Cantonal structure in Switzerland
- **bezirke.gpkg**: Dataset that depicts the Municipal structure in Switzerland

**R** and more specifically **sf package** offers options (functions) not only for 
transforming the coordinate system of a dataset, but also for identifying the
current existing one. Let's dive in :)

```{r echo=TRUE, message=TRUE, warning=TRUE}
# Importing the datasets
library(sf)
kantone <- read_sf("sample_data/other/kantone.gpkg")
bezirke <- read_sf("sample_data/other/bezirke.gpkg") 
```


```{r echo=TRUE, message=TRUE, warning=TRUE}
# Visualising the data - Cantons
library(ggplot2)
ggplot() + geom_sf(data = kantone)
```

```{r echo=TRUE, message=TRUE, warning=TRUE}
# Visualising the data - Bezirke
library(ggplot2)
ggplot() + geom_sf(data = bezirke)
```

In order to identify the current coordinate system of a dataset, **sf package**
offers the function `st_crs`. The main usage of the function is to retrieve th
coordinate reference system from an sf object. 

### Reproject raster data in R

















## Topology Toolset

## Sampling Toolset

### Generate Tessellation

- Hexagon
- Tranverse Hexagon
- Square
- Diamond
- Triangle


### Create Fishnet

### Create Random Points

### Generate Points Along Lines

### Generate Rectangles Along Lines

### Generate Transsects Along Lines

## Raster Toolset

<!-- Yes, these are toolsets within toolsets -->
### Mosaic Dataset Toolset

### Ortho Mapping Toolset

### Raster Dataset Toolset

### Raster Processing Toolset

#### Clip Raster

#### Composite Bands

#### Compute Pansharpen Weights

#### Create Ortho Corrected Raster Dataset

#### Create Pansharpened Raster Dataset

#### Extract Subdataset

#### Generate Table From Raster Function

#### Raster To DTED

#### Resample

#### Split Raster


### Raster Properties Toolset


