# Image Analysis Toolbox

<!-- ## Map Algebra Toolset -->


### Raster Calulator

In GIS Softwares (ArcGIS pro, QGIS etc) `Raster Calculator` tool is one of the most
commonly used tools for performing various operations with raster datasets.  
Let's explore our options for performing similar tasks with R.  

```{r}
library(raster)
pop_2000 <- raster("sample_data/population_raster/gpw_v4_population_count_rev11_2000_2pt5_min.tif")
pop_2010 <- raster("sample_data/population_raster/gpw_v4_population_count_rev11_2010_2pt5_min.tif")

par(mfrow=c(1,2))
plot(pop_2000)
plot(pop_2010)

```

<!-- ## Math Toolset -->

<!-- ## Overlay Toolset -->

<!-- ### Weighted Sum -->

<!-- ## Statistical Toolset -->

<!-- ### Cell Statistics -->

### Focal Statistics

>`Focal statistics` performs a neighbourhood operation that computes an output raster, where the value for each output cell is a function of the values of all the input cells that are in a specified neighborhood around that location. The function performed on the input is a statistic, such as the maximum, average, or sum of all values encountered in that neighborhood.
Source: https://pro.arcgis.com/en/pro-app/latest/tool-reference/spatial-analyst/how-focal-statistics-works.html

In R the respective operation can be performed with the use of the _focal_ function of the __terra package__.
As an example below, we will create our own raster object.  

```{r}
library(raster)
library(terra)

r <- raster(ncol=10, nrow=10)

ncell(r)
values(r) <- 1:ncell(r)
# # inMemory(r)
values(r)[1:10]
plot(r, main='Raster with 100 cells')

``` 
<!-- ### Zonal Statistics -->