## Importing raster data with `raster` or `terra`

```{r, include=FALSE}
source("R/before_chapter.R", local = knitr::knit_global())
```

To import raster data, we need a different packge. Currently, `raster` is the go-to package to deal with raster datasets. However, this package is currently in the process of being replaced with a successor, `terra`. We will showcase both methods here.



```{r}
library(raster)

dhm_200 <- raster("sample_data/Raster/dhm_200.tif")

dhm_200

```


```{r}
class(dhm_200)
```

`RaysterLayer`s also come with a plot method to visualize the dataset spatially:

```{r}
plot(dhm_200)
```



@todo add terra method to import rasters
@todo add method for rasters with multiple layers

```{r}


# path_1b <- "sample_data/SMR_Musterdaten/SMR50_LV95_KOMB_Mosaic.tif"
# path_3b <- "sample_data/SMR_Musterdaten/SMR50_LV95_KREL_10L_Mosaic.tif"
# 
# raster(path_1b)
# raster(path_3b)
```



As you can see in the output of `raster(path_3b)`, only one of the 3 bands was imported with `raster::raster`. In the case of multiband rasters, it's better to use `raster::brick`:

```{r}
# smr50_komb_mosaic <- raster(path_1b)
# smr50_krel_mosaic <- brick(path_3b)
```