---
output: github_document
---

```{r, include= FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


# All things `R`-Spatial


[![Actions Status](https://github.com/arc2r/book/workflows/bookdown/badge.svg)](https://github.com/arc2r/book/actions)


## Geoprocessing tools

```{r}
library(tidyverse)
library(data.tree)

tools <- read_csv("ESRI_Tool_names.csv")


treefy <- tools %>%
  mutate(tool = paste(toolb,toolset,Tool, sep = "/"))  %>% 
  as.Node(pathName = "tool")

print(treefy,limit = NULL)
```



## Style Guide

Some basic text formatting rules for keeping the document consistent.


### Inline formatting

- formatting with __bold__, in the following cases:
    - dataset name (e.g __parking spots in the canton of Zurich__)
    
- formatting with _italics_, in the following cases:
    - referring to a tool (e.g `Select By Attributes`) 
    - operation (e.g __spatial subsetting__)

    
- formatting with this `format`, in the following cases:
    - package name (e.g `sf`)
    - function name (e.g `st_buffer`)
    
### Adding notes

If you want to highlight something important, you can add [a custom block](https://bookdown.org/yihui/bookdown/custom-blocks.html) in this format: ````{block2, type = "rmdnote"}` (see example below). This block will recieve the class "rmdnote", which is specified in [css/style.css]. Currently, the classes `rmdcaution, .rmdimportant, .rmdnote, .rmdtip, .rmdwarning` have the same layout, they all come from "Geocomputation with R" and might be specified on demand.

````markdown
```{block2, type='rmdnote'}
Something very important we want to convey to our readers, written in Markdown.
```
````

### Subdividing into Parts

We can add Subdivision to our document by adding the following line at the top of an Rmd-File

```
# (PART) Topology Rules {-}
```

see https://bookdown.org/yihui/bookdown/markdown-extensions-by-bookdown.html#special-headers


### Todos:

we can add todos to our document using the following syntax:

```
/ todo: the title of the todo. what should be done?
/  here we can add a body of our todo, which is indeted with an additional space
/  we can also automatically assign a todo to a user using their github username
/  in the following manner:
/  asignees: ratnanil
```

The Todo action by alstr (https://github.com/alstr/todo-to-issue-action/) will automatically check all our files for new todos and create issues if it finds one. Once the todo is removed from the document, the issue will be closed as well.

If we want the todo to be invisible from the users, we need to wrap it with the html comment tag `<!--` and `-->` as follows. Note that the comment tags are written on separate lines.

```
<!--
/ todo: the title of the todo. what should be done?
/  here we can add a body of our todo, which is indeted with an additional space
/  we can also automatically assign a todo to a user using their github username
/  in the following manner:
/  asignees: ratnanil
-->
```


### Citation

To cite a book or URL or whatever, add a appropriate bibtex entry to [bibliography/general.bib]. You can then cite the document using the reference \@key  or [\@key].

[bibliography/packags.bib] is generated automatically from all the packages in the DESCRIPTION of the book and the DESCRIPTION of the repo arc2r/arc2r. For this reason, ALL items in these bibtex files are added to the document, so only add entries for items we use in the book!


    
## Dependencies

To capture the various dependencies on r-Packages, we follow the following logic (see also https://github.com/arc2r/book/issues/9): 



| Priority 	| Dependency (Package Necessary to:)            	| File        	| Repo        	| Field    	| Eg.                 	|
|----------	|-----------------------------------------------	|-------------	|-------------	|----------	|---------------------	|
| 1        	| Use the **data-package**                      	| DESCRIPTION 	| arc2r/arc2r 	| Imports  	| `sf`, `raster`          	|
| 2        	| To run the code described within **the book** 	| DESCRIPTION 	| arc2r/arc2r 	| Suggests 	| `gstat`, `tmap`, `dplyr`  	|
| 3        	| Render **the book**                           	| DESCRIPTION 	| arc2r/book  	| Imports  	| `bookdown`, `rmarkdown` 	|
