## Projections and Transformations Toolset {#projections}

For this chapter, you will need the following R Packages:

```{r, echo = TRUE}
library(arc2r)
library(sf)
library(raster)
```

If there is anything special about spatial, it's the Coordinate Reference System. We will not get into the topic of CRS and Projections here, but it is advisable to get up to speed on this topic if you are dealing with geodata. From this toolset, we will cover the following tools:

- **Define Projection**
- **Project**
- **Project Raster**
- **Rescale** (Resizes a raster by the specified x and y scale factors)


### Define Projection {#projections-defineprojection}

> Overwrites the coordinate system information (map projection and datum) stored with a dataset. This tool is intended for datasets that have an unknown or incorrect coordinate system defined.

If your dataset comes *without* an assigned CRS, things can get hairy. Usually, you can make an informed guess as to which CRS the dataset is associated with, but in the end only the data producer can tell you. Let's take the dataset `bezirke` from the package `arc2r`. As you can see, the `CRS` field in the data's header is `NA`. 

```{r, echo = TRUE}
data("bezirke")
```

Calling the function `st_crs()` on `bezirke` confirms this.

```{r, echo = TRUE}
st_crs(bezirke)
```

In ArcGIS, we would use the Tool "Define projection". In `R`, we use the function `st_crs(dataset) <- value`, where `value` is the CRS of our dataset. It can be either

1. A string accepted by GDAL (type `character`)
2. A valid EPSG value (type `integer`) *or*
3. An object of class crs

We find option two to be the simplest approach. 
<!--
todo how do we know this?
gloss over this?
-->
Since we know the dataset `bezirke` is in the new swiss coordinate system (CH1903+ / LV95), we can find out the EPSG code (e.g. by consulting [epsg.io](http://epsg.io/)).
We will not get into options 1 and 3 and only 

```{r, echo = TRUE}
st_crs(bezirke) <- 2056
```

If we look at our header again, we see that we have a new field `projected CRS` which is in fact displaying the correct name based on the EPSG code we provided. 

```{r, echo = TRUE}
bezirke
```

```{block2, type = "rmdnote", echo = TRUE, purl = FALSE}
<!--
# TODO revise this
-->
 _st_set_crs_ function **does not** reproject the coordinates of the given dataset. In other words, it does not affect the actual _geometry_ column of the sf object. _st_tranform_ on the other hand indeed **does indeed reproject** the dataset to another coordinate
```


### Project {#projections-project}

> Projects spatial data from one coordinate system to another

Now that `bezirke` has an assigned `CRS` (see \@ref(projections-defineprojection), we can transform it into a new coordinate system. To emphasize this, will will visualize the dataset using `base::plot()` with the option `axes = TRUE`, to visualize the current coordinate system.  

```{r echo=TRUE, message=TRUE, warning=TRUE}
plot(bezirke["area_km2"], axes = TRUE)
```

```{r echo=TRUE, message=TRUE, warning=TRUE}
bezirke_wgs84 <- st_transform(bezirke, 4326)

plot(bezirke_wgs84["area_km2"], axes = TRUE)
```

```{r echo=TRUE, message=TRUE, warning=TRUE}
st_crs(bezirke)
```




```{r echo=TRUE, message=TRUE, warning=TRUE}
bezirke_swiss <- st_transform(bezirke, 2056)

# retrieve the coordinate system
st_crs(bezirke_swiss)
```


### Project Raster {#projections-projectraster}

> Transforms a raster dataset from one coordinate system to another

Working with Raster datasets in GIS of operations is of equal importance, as
working with vector ones. One of the spatial properties of raster datasets is the 
the Coordinate Reference System (CRS). CRS is the specific system that “associates”
the raster coordinates (which are just pairs of x/y values) to geographic locations.
In ArcGIS pro the tool for projecting a raster dataset is called
_Project Raster (Data Management)_. Let's see how we can perform the same operation
with R. 


```{r, echo = TRUE}
# Dataset derived from the spatial interpolation of all the available "recycling points"
# in the city of Wädenwil
data("recycling_raster")

# Dataset representing the public transport quality in the city of Wädenswil
data("public_transport_waedi")# CRS -> WGS84 
```


```{r, echo = TRUE}
# Plot the raster dataset - World Geodetic System 1984 
plot(public_transport_waedi,las=1,
     main = "Quality of public transport in the city of Wädenwil - CRS: WGS84",
     cex.main=1,font.main=4)
```

We can use the `projectRaster()` function to reproject a raster into a new CRS.
The first argument of the aforementioned function is the raster dataset we want to reproject,
while the second one is the dataset to whose projection we are targeting to. So, in
our case, we are targeting to the coordinate system of the __raster_recycling__
dataset.  
It is important to remember that raster reprojection only works when the raster
object has already a defined CRS.

```{r, echo = TRUE}
# Transform the coordinate system of the raster dataset publicTransport_CH
# into the Swiss Coordinate system - CH1903+LV95
publicTransport_CH = projectRaster(public_transport_waedi, recycling_raster)
```

```{r, echo = TRUE}
# Plot the raster dataset - Swiss Coordinate System CH1903+LV95 
plot(publicTransport_CH,las=1,
     main = "Quality of public transport in the city of Wädenwil - CRS: CH1903+LV95",
     cex.main=1,font.main=4)
```