# Image Analysis Toolbox

For this chapter, you will need the following R Packages:
```{r, echo=FALSE}
library(arc2r)
library(raster)
library(terra)
```

## Map Algebra Toolset



### Raster Calulator

In GIS Softwares (ArcGIS pro, QGIS etc) `Raster Calculator` is one of the most
commonly used tools for performing various operations with raster datasets.  

For the example below we will work with 2 datasets that depict India's population at two different timestamps. Our goal is to produce an output raster that demonstrates the evolvement of country's population in this 10 years timespan. Naturally in ArcGIS pro, the aforementioned operation would be performed with the use of `Raster Calculator`. Let's explore our options with R.  

<!--
/ todo: replcace this with arc2r/arc2r
/  assignees: KimboGilda
--> 

```{r}
# Read the data for our example

# India's population in 2000
data("pop_india2000")

# India's population in 2010
data("pop_india2010")

par(mfrow=c(1,2))
plot(pop_india2000)
plot(pop_india2010)
```

Let's compute now the difference between the two rasters. To do so, there are various 
options. As first we can create a function that performs a simple subtraction of two
rasters.


```{r}
diff_rasters <- function(b1, b2){
  # this function calculates the difference between two rasters of the same CRS and extent
  # input: 2 raster layers of the same extent, crs that can be subtracted
  # output: a single different raster of the same extent, crs of the input rasters
  diff <- b2 - b1
  return(diff)
}
```

Since, we created the function for subtracting one raster from another, we use now
the function _overlay_ of the `raster` package for producing the new raster output.  

```{r}
# Compute the raster that depicts the difference in population in these 10 years
pop2010_2000 <- overlay(pop_india2000, pop_india2010,
                     fun = diff_rasters)

plot(pop2010_2000,
     main = "India's population difference in the timespan of 10 years",
     cex.main = 1)
```


<!-- ## Math Toolset -->

<!-- ## Overlay Toolset -->

<!-- ### Weighted Sum -->

<!-- ## Statistical Toolset -->

<!-- ### Cell Statistics -->

### Focal Statistics

### Zonal Statistics

>`Focal statistics` performs a neighbourhood operation that computes an output raster, where the value for each output cell is a function of the values of all the input cells that are in a specified neighborhood around that location. The function performed on the input is a statistic, such as the maximum, average, or sum of all values encountered in that neighborhood.
Source: https://pro.arcgis.com/en/pro-app/latest/tool-reference/spatial-analyst/how-focal-statistics-works.html

In R in order to apply functions focally to rasters, important is to define the neighbouring cells.
In other words to define the moving window for each of these functions. As an example below,
we will create our own raster object.  

```{r}
r <- raster(ncol=10, nrow=10)
values(r) <- 1:ncell(r)
plot(r, main='Raster with 100 cells', cex.main = 1)
``` 

We will recompute now the raster based on the `focal` function of the  `raster` package. In the function
below, the neighbourhood for which we apply the statistic is represented by the argument `w`, while the 
statistic by the argument `fun`. 

```{r} 
f1 <- focal(r, w=matrix(1,nrow=3,ncol=3),fun=sum)
plot(f1, main = "Raster after applying the focal function - raster package", cex.main = 1)
```

The operation above can be also performed with the use of the _focal_ function of the __terra__ package.  

  - `w` represents the moving window of the function. The window can be defined as one (for a square)
    or two numbers (row, col); or with an odd-sized weights matrix. More details can be found here: https://rdrr.io/cran/terra/man/focal.html
  - `fun` represents the specified statistic for the identified neighbourhood (e.g mean,sum etc)
  
```{r}

r_spat <- as(r, "SpatRaster")
f2 <- terra::focal(r_spat,w=3,fun = "sum")
plot(f2, main = "Raster after applying the focal function - terra package", cex.main = 1)
```

### Zonal Statistics

`Zonal Statistics` tool in ArcGIS pro is used for calculating the statistics
on values of a raster within the zones of another dataset. In R we are able to produce
similar results with the use of `zonal` function of the `raster` package. 

```{r}
r <- raster(ncols=10, nrows=10)
# generating random numbers that will used as values in our raster layer (r) - input values raster
values(r) <- runif(ncell(r)) * 1:ncell(r)

z <- r
# generating integer values for the raster layer z, which will be used as the zone raster
values(z) <- rep(1:5, each=20) 

# Zonal statistics - Summarize
zonal(r, z, "sum")

# Zonal statistics - Mean
zonal(r, z, "mean")

# Zonal statistics - Minimum
zonal(r, z, "min")
```



