```{r}
library(sf)
library(tidyverse)
```


# Spatial Analyst Toolbox

## Conditional Toolset
## Density Toolset

### Kernel Denisity


There are several Function that can be tweaked to calculate KDE for `sf`-Point object:

- `tmaptools::smooth_map()`: Depricated (is there a successor?)
- `spatstat::density.ppp()`: Takes only objects of Class `ppp`
- `MASS::kde2d()`: Takes x/y coordinates as vectors and returns a matrix

In this example, I take `MASS:kde2d()` and tweak it to take `sf` and return `raster`. First, let's create some sample data:

```{r}
set.seed(10)
mypoints <- data.frame(x = rnorm(1000),y = rnorm(1000)) %>%
  st_as_sf(coords = c(1,2))

plot(mypoints)
```


```{r}
my_kde <- function(points,cellsize, bandwith, extent = NULL){
  require(MASS)
  require(raster)
  require(sf)
  if(is.null(extent)){
    extent_vec <- st_bbox(points)[c(1,3,2,4)]
  } else{
    extent_vec <- st_bbox(extent)[c(1,3,2,4)]
  }
  
  n_y <- ceiling((extent_vec[4]-extent_vec[3])/cellsize)
  n_x <- ceiling((extent_vec[2]-extent_vec[1])/cellsize)
  
  extent_vec[2] <- extent_vec[1]+(n_x*cellsize)-cellsize
  extent_vec[4] <- extent_vec[3]+(n_y*cellsize)-cellsize

  coords <- st_coordinates(points)
  matrix <- kde2d(coords[,1],coords[,2],h = bandwith,n = c(n_x,n_y),lims = extent_vec)
  raster(matrix)
}
```

```{r}

mypoints_kde <- my_kde(mypoints,0.01,1)
library(stars)

ggplot() +
  geom_stars(data = st_as_stars(mypoints_kde)) +
  geom_sf(data = mypoints, alpha = 0.2, fill = "black") +
  scale_fill_viridis_c() +
  labs(fill = "KDE") +
  theme_void() 


```


### Line Density

### Point Density

## Distance Toolset

## Extraction Toolset

## Generalization Toolset

## Interpolation Toolset

## Map Algebra Toolset

## Math Toolset

## Neighborhood Toolset

## Overlay Toolset

### Fuzzy Membership
### Fuzzy Overlay
### Locate Regions
### Weighted Overlay
### Weighted Sum


## Reclass Toolset

### Lookup
### Reclass by ASCII File
### Reclass by Table
### Reclassify
### Rescale by Function
### Slice


## Surface Toolset

### Aspect
### Contour
### Contour List
### Contour with Barriers
### Curvature
### Cut Fill
### Hillshade
### Observer Points
### Slope
### Viewshed
### Viewshed 2
### Visibility


## Zonal Toolset






