## Spatial intersections

### Clip operation
```{r,echo=FALSE}
knitr::opts_chunk$set(warning = F, message = F, tidy = T)
```


```{r}
bahn_haltestelle_ch <- st_read("sample_data/other/Haltestelle_Bahn_CH.shp")
kantonsgebiet <- st_read("sample_data/other/Kantonsgebiet.shp")

tm_shape(kantonsgebiet) + tm_polygons("#f0f0f0") + 
  tm_shape(bahn_haltestelle_ch) + tm_dots(col = "#e34a33",size = 0.1, alpha = 0.5) + 
  tm_shape(kantonsgebiet) + tm_borders(col = "black") + tm_layout(frame = F)
   
```
We want now to analyse the situation at a particular area. So we are going to create an index with the specific cantons we are interested in. 
```{r}
index <- kantonsgebiet$name == "Zürich" |
  kantonsgebiet$name == "St. Gallen" |
  kantonsgebiet$name == "Thurgau" |
  kantonsgebiet$name == "Aargau"

kantons_aOi <- kantonsgebiet[index,]

hal_clipped <- st_intersection(kantons_aOi,bahn_haltestelle_ch)
```
Plot the result from the clipping operation
```{r}
tm_shape(kantons_aOi) + tm_polygons("#f0f0f0") + 
  tm_shape(hal_clipped) + tm_dots(col = "#e34a33",size = 0.1, alpha = 0.5) + 
  tm_shape(kantons_aOi) + tm_borders() + tm_layout(frame = F)
```


So, ultimately, as we can see above, the _st_intersection_ function creates a datasets where the point dataset is precisely "clipped" based on the area of interest. 