## Spatial Intersections

### Clip operation
```{r,echo=FALSE}
knitr::opts_chunk$set(warning = F, message = F, tidy = T)
```


```{r}
bahn_haltestelle_ch <- st_read("sample_data/other/Haltestelle_Bahn_CH.shp")
kantonsgebiet <- st_read("sample_data/other/Kantonsgebiet.shp")

tm_shape(kantonsgebiet) + tm_polygons("#f0f0f0") + 
  tm_shape(bahn_haltestelle_ch) + tm_dots(col = "#e34a33",size = 0.1, alpha = 0.5) + 
  tm_shape(kantonsgebiet) + tm_borders(col = "black") + tm_layout(frame = F)
   
```
We want now to analyse the situation at a particular area. So we are going to create an index with the specific cantons we are interested in. 
```{r}
index <- kantonsgebiet$name == "Zürich" |
  kantonsgebiet$name == "St. Gallen" |
  kantonsgebiet$name == "Thurgau" |
  kantonsgebiet$name == "Aargau"

kantons_aOi <- kantonsgebiet[index,]

hal_clipped <- st_intersection(kantons_aOi,bahn_haltestelle_ch)
```
Plot the result from the clipping operation
```{r}
tm_shape(kantons_aOi) + tm_polygons("#f0f0f0") + 
  tm_shape(hal_clipped) + tm_dots(col = "#e34a33",size = 0.1, alpha = 0.5) + 
  tm_shape(kantons_aOi) + tm_borders() + tm_layout(frame = F)
```
So, ultimately, as we can see above, the _st_intersection_ function creates a result where the point dataset is precisely "clipped" based on the area of interest.

### Buffer operation

One of the most commonly used operations in ArcGIS pro is the one called `Buffer`. This operation provides a very convenient way of identifying areas of interest lying in the neighborhood of an existing spatial feature. In R the same action can be performed using the _st_buffer_ function in _sf_. 

Below we introduce a line spatial feature that depicts all the __mountain bike routes__ in Switzerland.
```{r, echo=F}
mountainBikes_routes <- st_read("sample_data/other/mountainBikes_routes.shp")
```

```{r, echo=F}
tm_shape(kantonsgebiet) + tm_polygons("#f0f0f0") + 
  tm_shape(mountainBikes_routes) + tm_lines(col = "#e34a33") + 
  tm_shape(kantonsgebiet) + tm_borders(col = "black") + tm_layout(frame = F)
```

We focus, as an area of interest, on the canton of Valais. Consequently, we might be interested in identifying all the areas in a radius of 1000 m around these bike routes.

Selecting the canton of Valais
```{r, echo=FALSE}
index2 <- kantonsgebiet$name == "Valais"
canton_valais <- kantonsgebiet[index2,]
mountainBikes_routes_valais <- st_intersection(mountainBikes_routes,canton_valais)

tm_shape(canton_valais) + tm_polygons("#f0f0f0") + 
  tm_shape(mountainBikes_routes_valais) + tm_lines(col = "#e34a33") + 
  tm_shape(canton_valais) + tm_borders(col = "black") + tm_layout(frame = F)
```

Apply a buffer of 1000m on each of the mountain bike routes within the canton of Valais
```{r}
routes_buffer <- st_buffer(mountainBikes_routes_valais,1000)

tm_shape(canton_valais) + tm_polygons() + 
  tm_shape(mountainBikes_routes_valais) + tm_lines(col = "#e34a33") +
  tm_shape(routes_buffer) + tm_polygons(col = "blue", alpha = 0.2) + 
  tm_shape(canton_valais) + tm_borders(col = "black") + tm_layout(frame = F)
```
