# Spatial Predicates {#spatial-predicates}

In Opensource GIS Applications, there is a common standard to describe various relationships between features. They are referred to as ["Spatial predicates"](https://en.wikipedia.org/wiki/DE-9IM#Spatial_predicates) and are defined as follows. Take two geometries, `a` and `b`:

- **Equals**: `a` and `b` are equal if
  - their interiors intersect and
  - no part of the interior or boundary of one geometry intersects the exterior of the other
- **Disjoint**: `a` and `b` are disjoint if they have no point in common (they form a set of disconnected geometries)
- **Touches**: `a` and `b` touch if
  - they have at least one point in common but
  - their interiors do not intersect
- **Contains**: `a` contains `b` if
  - no points of `b` lie in the exterior of `a` and
  - at least one point of the interior of `b` lies in the interior of `a`
- **Covers**: `a` covers `b` if at least one point of `b` lies in `a`, and no point of `b` lies in the exterior of `a`
- **Intersects**: `a` and `b` intersect if the two geometries have at least one point in common
- **Within** (inside): `a` is within `b` if `a` lies in the interior of the `b`
- **CoveredBy**: `a` is covered by `b` if `a` lies in `b`
- **Crosses**: `a` crosses `b` if they have some but not all interior points in common.
- **Overlaps**: `a` overlaps `b`
  - if they have some (but not all) points in common and
  - they have the same dimension and
  - the intersection of the interiors of the two geometries has the same dimension as the geometries themselves



## Named predicates

### Touches

Lets take the example of a chessboard:

```{r, echo = TRUE}
library(sf)
library(dplyr)
library(arc2r)

data("chessboard")


chessplot <- ggplot(chessboard) + 
  geom_sf(aes(fill = colour)) + 
  geom_sf_label(aes(label = i)) +
  scale_x_continuous(breaks = 1:8-0.5,labels = letters[1:8]) +
  scale_y_continuous(breaks = 1:8-0.5,labels = 1:8) +
  scale_fill_manual(values = c(black = "black",white = "white"))+ 
  theme_void() +
  theme(legend.position = "none") 

chessplot
```

To find out which field `touch` field number 36, we can write the following line of code:

```{r, echo = TRUE}
st_touches(chessboard[36,],chessboard)
```

Visually, these are the following fields:

```{r, echo = TRUE, purl = FALSE}
sel36 <- st_touches(chessboard[36,],chessboard)[[1]]


chessplot +
  geom_sf(data = chessboard[36,], fill = "blue", alpha = 0.4)+
  geom_sf(data = chessboard[sel36,], fill = "red",alpha = 0.4)
```