## Surface Toolset

### Slope and Aspect

>In mathematics, the slope or gradient of a line describes its steepness, incline, or grade. A higher slope value indicates a steeper incl.  
Source: "http://wiki.gis.com/wiki/index.php/Slope"

In GIS and specifically in terrain analysis, calculating the terrain _slope_ is of
great importance, since it can play a significant role in various forms of technical
analysis. The most common way of calculating the _slope_ is from a Digital Elevation Model (DEM).  
  
In ArcGIS pro the tool used for computing the slope of a raster dataset with elevation
data is called `Slope (3D Analyst)`.It identifies the steepness at each cell of a raster surface.
The lower the slope value, the flatter the terrain; the higher the slope value, the steeper the terrain.  
  
In `R`, `raster` package includes the function `terrain()`, which helps us compute
*slope*, *aspect* and other terrain characteristics from a raster with elevation data.  
Let's start by computing the _slope_ of the **swissAlti3d** raster dataset.  

<!--
/ todo: replcace this with arc2r/arc2r
/  assignees: KimboGilda
--> 

```{r, echo = TRUE, fig.cap = "swissALTI3D2019 - CRS: CH1903+LV95"}
data("swissAlti3D")

plot(swissAlti3D)
```

After importing and visualizing our raster dataset, we use the `terrain` function
to calculate the slope.  

```{r, fig.cap = "swissALTI3D2019 - CRS: CH1903+LV95"}
# Computing the slope of a raster dataset using the terrain() function
swissAlti3d_slope <- terrain(swissAlti3D, opt="slope", unit="degrees")
plot(swissAlti3d_slope)
```
  
In the function above, we define the operation we want to perform as the second
argument of the function (_opt = "slope"_). Furthermore, we also define the units
of the final result.  
Similarly, we compute the `aspect` of a given dataset. Basically as `aspect` we can
consider the compass direction that a `slope` leans towards.  
Let's compute now, using again the same function `terrain`, the `aspect` of the
same dataset __swissALTI3D2019__.   

```{r, fig.cap="Aspect of swissALTI3D2019 in degrees (°)"}
# Computing the aspect of the terrain dataset
swissAlti3d_aspect <- terrain(swissAlti3D, opt="aspect", unit="degrees")
plot(swissAlti3d_aspect)
```  
  
So, R gives us the option to retrieve multiple terrain characteristics of a raster
dataset, using only one function. The aforementioned spatial operation in ArcGIS pro
could be performed using the tool `Aspect (Spatial Analyst)`.

