## Surface Toolset

### Slope and Aspect

>In mathematics, the slope or gradient of a line describes its steepness, incline, or grade. A higher slope value indicates a steeper incl.  
Source: "http://wiki.gis.com/wiki/index.php/Slope"

In GIS and specifically in terrain analysis, calculating the terrain _slope_ is of
great importance, since it can play a significant role in various forms of technical
analysis. The most common way of calculating the _slope_ is from a Digital Elevation Model (DEM).  
  
In ArcGIS pro the tool used for computing the slope of a raster dataset with elevation
data is called `Slope (3D Analyst)`.It identifies the steepness at each cell of a raster surface.
The lower the slope value, the flatter the terrain; the higher the slope value, the steeper the terrain.  
  
In `R`, `raster` package includes the function `terrain()`, which helps us compute
*slope*, *aspect* and other terrain characteristics from a raster with elevation data.  
Let's start by computing the _slope_ of the **swissAlti3d** raster dataset.  

<!--
/ todo: replcace this with arc2r/arc2r
/  assignees: KimboGilda
--> 

```{r}
# Importing a raster dataset
# swissALTI3D2019.tif -- Source: https://geovite.ethz.ch
data("swissAlti3D")

plot(swissAlti3D,
     main = "swissALTI3D2019 - CRS: CH1903+LV95",
     cex.main=1,font.main=4)
```

After importing and visualizing our raster dataset, we use the `terrain` function
to calculate the slope.  

```{r}
# Computing the slope of a raster dataset using the terrain() function
swissAlti3d_slope <- terrain(swissAlti3D, opt="slope", unit="degrees")
plot(swissAlti3d_slope,
     main = "Slope of swissALTI3D2019 in degrees (°)",
     cex.main=1,font.main=4)
```
  
In the function above, we define the operation we want to perform as the second
argument of the function (_opt = "slope"_). Furthermore, we also define the units
of the final result.  
Similarly, we compute the `aspect` of a given dataset. Basically as `aspect` we can
consider the compass direction that a `slope` leans towards.  
Let's compute now, using again the same function `terrain`, the `aspect` of the
same dataset __swissALTI3D2019__.   

```{r}
# Computing the aspect of the terrain dataset
swissAlti3d_aspect <- terrain(swissAlti3D, opt="aspect", unit="degrees")
plot(swissAlti3d_aspect,
     main = "Aspect of swissALTI3D2019 in degrees (°)",
     cex.main=1,font.main=4)
```  
  
So, R gives us the option to retrieve multiple terrain characteristics of a raster
dataset, using only one function. The aforementioned spatial operation in ArcGIS pro
could be performed using the tool `Aspect (Spatial Analyst)`.


### Reclassify

In GIS quite often arises the necessity to reclassify a raster dataset. In other
words, to create new classes with different range of values for the existing cell
values of the dataset. This operation in ArcGIS Pro is performed using the
`Reclassify (Spatial Analyst)` tool.  
In R the respective operation is quite straightforward and is based on the use of
the `reclassify` function of the `raster` package. For our example we use the
raster dataset we computed previously, which depicts the `aspect` of the __swissALTI3D2019__
dataset.  

As a first step, we can compute the histogram of our dataset.  

```{r}
summary(swissAlti3d_aspect)
histinfo <- hist(swissAlti3d_aspect)

histinfo$breaks
```

The number of breaks for our raster values are 19. The aim of ours is to create 4
different classes with the following values:  

- 0°-90° -> Class 1
- 90°-180° -> Class 2
- 180°-270° -> Class 3
- 270°-360° -> Class 4  
  
To do so, we create below a reclassification matrix with the respective values.

```{r}
# create a reclassification matrix
reclass_df <- c(0,90,1,90,180,2,180,270,3,270,360,4)

# convert to Matrix
reclass_m <- matrix(reclass_df,
                ncol = 3,
                byrow = TRUE)

reclass_m
```

After creating the respective matrix with the new classes, we are ready to
call the `reclassify` function. The function takes as a first argument the dataset
set to be reclassified and as second the matrix, on which the reclassification is
based on. 

```{r}
swissAlti3d_aspect_reclass <- reclassify(swissAlti3d_aspect,
                     reclass_m)

plot(swissAlti3d_aspect_reclass, col = c("red", "green", "yellow","blue"),
     main = "Aspect of swissALTI3D2019 in 4 classes",
     cex.main=1,font.main=4, legend = FALSE)

legend("topright",
       legend = c("0°-90°", "90°-180°", "180°-270°","270°-360°"),
       fill = c("red", "blue", "green", "yellow"),
       border = FALSE,
       bty = "n") # turn off legend border
```






<!-- ### Contour -->
<!-- ### Contour List -->
<!-- ### Contour with Barriers -->
<!-- ### Curvature -->
<!-- ### Cut Fill -->
<!-- ### Hillshade -->
<!-- ### Observer Points -->
<!-- ### Viewshed -->
<!-- ### Viewshed 2 -->
<!-- ### Visibility -->