---
output: github_document
---

```{r}
library(tidyverse)
library(data.tree)
library(networkD3)
library(yaml)

knitr::opts_chunk$set(echo = TRUE)

```

This document has two tasks. It serves as a reference to show how we document the "Geoprocessing Tools" we want to cover in the book (a yaml file). Additionally, it creates the network diagrams.

This way to document the tools has the following advantages:
1. Tools can be uniquely identified using Toolbox and Toolset
2. It allows minimal repetition, since Toolboxes and Toolsets must be noted only once.


The yaml file is structured as follows:

```{yaml}
--- #  Geoprocessing Tools    # This is the yaml header and  a comment to show what this is about
  Toolbox:                    # Level 1 (indented 2 spaces): The toolbox's name                
    Toolset:                  # Level 2 (indented 4 spaces): The toolset's name 
      - Tool 1                # Level 3 (indented 6 spaces): the tool's name (starts with a dash)
      - Tool 2                # Level 3 (indented 2 spaces): the tool's name (starts with a dash)

```


```{r}
tools_toodo <- read_yaml("auxiliary_files/tools_todo.yaml")
```

This next line of code converts the tools (which are imported as a character vector) into a list.

```{r}
tools_toodo2 <- map(tools_toodo_yaml, function(x){map(x, function(y){map(y, function(z){list(name = z)})})})
```


```{r}

FromListSimple(tools_toodo2,nodeName =  "GeoprocessingTools") %>%
    ToListExplicit(unname = TRUE) %>%
    diagonalNetwork()

imap(tools_toodo2, function(toolsets,toolbox){
  FromListSimple(toolsets,nodeName = toolbox) %>%
    ToListExplicit(unname = TRUE) %>%
    diagonalNetwork() %>%
    networkD3::saveNetwork("network.html")
  
  webshot::webshot("network.html",file.path("images","network",paste0(toolbox,".png")))
  file.remove("network.html")
  imap(toolsets,function(tools,toolset){
    FromListSimple(tools,nodeName = toolset) %>%
      ToListExplicit(unname = TRUE) %>%
      diagonalNetwork() %>%
      networkD3::saveNetwork("network.html")
    
    webshot::webshot("network.html",file.path("images","network",paste0(toolbox,"_",toolset,".png")))
    file.remove("network.html")
  })
})



```

