## Interactive Maps

As introduced in chapter [Making Maps], an interactive Map is the fundamental building block of ArcGIS. In R, we need to do some extra work to create such an interactive map, but this extra work sometimes means just one line of code. There are multiple packages that simplify creating maps:

- [`leaflet`](https://rstudio.github.io/leaflet/)
- [`mapview`](https://r-spatial.github.io/mapview/)
- [`tmap`](https://github.com/mtennekes/tmap)
- [`plotly`](https://plotly.com/r/)

`mapview` and `tmap` both build on the package `leaflet`, which in turn builds on [leaflet](https://leafletjs.com/), an open source JavaScript Library for interactive maps. There really is not much need to learn `leaflet`, since using `mapview` or `tmap` is much simpler. 

The package `plotly` builds on [plotly](https://plotly.com/), a very versatile graphics library that provides an API for `R`, `Python` and `Julia`. It is an amazing tool, but we will focus more on `mapview` and `tmap`.



### Package `mapview`

The package `mapview` is highly automated and you get a very decent map when calling the function `mapview::mapview` on your geodata, without any additional arguments. The aim of the package is to be quick and easy, as emphasized in the documentation:

> It’s main goal is to fill the gap of quick [...] interactive plotting to examine and visually investigate both aspects of spatial data, the geometries and their attributes.


```{r, echo=FALSE}
library(arc2r)
library(sf)
library(raster)
library(mapview)
```

This next line of code generates a representation of your data (in this case, `gemeindegrenzen_zh`), much like the default map in ArcGIS. 

```{r}
mapview(gemeindegrenzen_zh)
```

If the emphasis lies on creating a highly customized map, then `tmap` may be more appropriate.

### Package `tmap`

Like mentioned in the previous chapter, the package `tmap` facilitates creating highly customized maps. But the greatest strength in `tmap` comes from it's versatility: All you need to do is specify how you want your data to be displayed, and `tmap` can render this into an interactive map of a static one that you can print. All you need to do is change the `mode` within the function `tmap_mode()` to `view`.

```{r}
library(tmap)

tmap_mode("view")
```

As shown in chapter \@ref(static-maps-tmap), the `tmap` syntax very similar to `ggplot2`s, in the sense that connect "layers" using a `+` sign. 

```{r}
tm_shape(gemeindegrenzen_zh) +
  tm_polygons()
```

This interactive map is very similar to the one generated by `mapview`, which is not surprising since they are both built on top of `leaflet`.


### Further ressources

For more information on `tmap`, check out the package's gihub site: https://github.com/mtennekes/tmap
For more information on interactive maps in general, head over to the great book "Geocomputation with R": https://geocompr.robinlovelace.net/adv-map.html#interactive-maps

