# Making Interactive Maps

Probabbly one of the first things you want to do when working with a dataset is visualizing it spatially. This is basically always the first thing you do in ArcGIS. So how do you this in R?

If your aim is to explore the data, to just get a feel for it, your map needs to:

1. allow pan and zoom
2. display attribute information in popups
3. provide a background map
4. be easy to make

If all of these features are important to you, check out the next section on `tmap`. If interactivity (feature 1 & 2) aren't important to you, and especially if you are a season't `ggplot2` user, check the section on `ggplot2`. <!-- Add crossreference -->

There are many packages that provide these features. In terms of simplicity, versatility and scaleablity `tmap` is one of the best in my opion. To see how it works, import the usual libraries and the North Carlolina Sample data.

```{r}
library(sf)
library(tidyverse)
library(tmap)

nc <- st_read(system.file("shape/nc.shp", package = "sf"),quiet = TRUE)
```

Here's how the data looks in base plot:

```{r, echo = FALSE}
plot(st_geometry(nc))
```

`tmap` has two modes: a static mode for printable maps (called `plot`) and a dynamic mode for exploratory, interactive maps (called `view`). We want the latter for now, so we switch to the dynamic mode with: 
```{r}
tmap_mode("view")
```

`tmap` has based it's syntax loosely on `ggplot2`: Mainly, layer are added with `+` and aesthetics are specified within each layer.
```{r}
tm_shape(nc) +
  tm_polygons()
```

You can have a look at the attributes of each feature by clicking on it. To view the entire attribute table, you can print `nc` to the console, or have a look at it with `View(nc)` if you are using Rstudio.

Check the following ressources for more information:

- https://github.com/mtennekes/tmap
- https://geocompr.robinlovelace.net/adv-map.html#interactive-maps

