## Making Static Maps

Before you start making maps however, you will have to familiarize yourself with importing raster or vector data into R in the first place. For Vector data, you can read chapter [2.2 Vector data](https://geocompr.robinlovelace.net/spatial-class.html#vector-data) from the book "Geocomputation with R". The next chapter [2.3 Raster data](https://geocompr.robinlovelace.net/spatial-class.html#raster-data) provides a very good overview on rasterdata in R


### Static maps with `ggplot2`

#### Basics
```{r}
library(sf)
library(tidyverse)


getwd()
eire <- read_sf("sample_data/spData/eire.shp")
```

The most basic plot consists of a `ggplot()` initiation and a single layer. In case of `sf` data, you can add a layer with `geom_sf()`. If the coordinate system is unknown, the axes will be labelled with the numerical values of the coorindates.
```{r}
st_crs(eire)
ggplot(eire) + 
  geom_sf()
```

However, if the coordinate system *is* known, the WGS84 Coordinates are labelled and used as the major and minor grid lines.

```{r}
eire <- st_set_crs(eire, "+proj=utm +zone=30 +ellps=airy +units=km")

ggplot(eire) + 
  geom_sf()
```

As always in `ggplot`, if you want to use a column from your dataset as an input for, say, colour, you need to specify this with `aes()`:
```{r}
ggplot(eire) + 
  geom_sf(aes(fill = INCOME))
```

There are various ways to beautify this simple plot:

```{r}
ggplot(eire) + 
  geom_sf(aes(fill = log(INCOME)), colour = "grey") +
  labs(title = "Counties of the Irish Republic", 
       subtitle = expression("Distribution of Income (log"[10]*")"),
       fill = "") +
  scale_fill_viridis_c() +
  theme_void() +
  theme(legend.position = "bottom")
```

#### Adding Raster Maps Files (geotiffs)

While vector data is very easily implemented in ggplot2, rasterdata is a little harder. Its complicated enough with standard Raster Data, and even more so with raster maps.  There are two packages that facilitate the combination of rastermaps and ggplot: 

- [`rasterVis`](https://oscarperpinan.github.io/rastervis/)
- [`RStoolbox`](https://bleutner.github.io/RStoolbox/)

However, we will do without these packages to learn more on how rasters work and be as flexible as possible.

If you don't have the time or patience to read through the explanation, the following function takes a path to a raster and returns a dataframe with x/y coordinates as well as a column with the hex colours for the according cell.

```{r}
rastermap_to_dataframe <- function(path) {
  require(tiff)
  require(raster)
  raster_file <- raster::brick(path)
  bands_n <- nbands(raster_file)
  if(bands_n == 1){
    raster_indexed <- readTIFF(path,indexed = TRUE)
    raster_colormap <- attr(raster_indexed,"color.map")
    raster_colormap_hex <- rgb(raster_colormap[1,],raster_colormap[2,],raster_colormap[3,])
    raster_df <- data.frame(coordinates(raster_file), values = raster_colormap_hex[getValues(raster_file)+1])
  } else if(bands_n == 3){
    raster_df <- data.frame(
      coordinates(raster_file),
      values = rgb(values(raster_file[[1]]),values(raster_file[[2]]),values(raster_file[[3]]),maxColorValue = 255)
    )
  }
  return(raster_df)
}
```

##### With Colours

Tif files with colours can either be single band or include multiple bands in a single tif file. These need to be treated differently. Let's take two variants of the [swiss raster map 1:50k](https://shop.swisstopo.admin.ch/de/products/maps/digital_maps/digital/srm10) map. One is a single band dataset, the other a 3-band dataset.

```{r}
library(raster)
library(rdgal)

path_1b <- "sample_data/SMR_Musterdaten/SMR50_LV95_KOMB_Mosaic.tif"
path_3b <- "sample_data/SMR_Musterdaten/SMR50_LV95_KREL_10L_Mosaic.tif"

raster(path_1b)
raster(path_3b)
```

As you can see in the output of `raster(path_3b)`, only one of the 3 bands was imported with `raster::raster`. In the case of multiband rasters, it's better to use `raster::brick`:

```{r}
map_1b <- raster(path_1b)
map_3b <- brick(path_3b)
```

###### Singleband

A singleband RasterLayer can be regarded as vector of numeric values arranged in a matrix with some additional metadata. The metadata can be view with `str()`, the values can be obtained with `getValues()`.

```{r}
str(map_1b)
head(getValues(map_1b),50)

extent(map_1b)
```

The first 50 cells of this `RasterLayer` contain the value `37`. This value `37` is associated with a specific colour, which can be viewd in the `colortable` of the `RasterLayer`.

```{r}
coltab <- colortable(map_1b)
head(coltab)
scales::show_col(coltab,labels = FALSE)
```


In order to plot this `RasterLayer` in ggplot, we need to convert it into a dataframe. You can imagine that this is converting the wide matrix into a very long table with three columns: x and y coordinates of the lower left corner of each cell and a value specifying what the cell holds. 

```{r}
map_1b_df <- data.frame(coordinates(map_1b),values = getValues(map_1b))
head(map_1b_df)
```

We can now use the `values` column in `ggplot2` to specify the `fill` aesthetic. If we then pass the colortable as our fill values, the colours will be matched correctly. However we need two additional tricks:

- name the colourtable to make sure the correct colours are correct to the according values
- wrap the `values` column in `factor()` so that the `ggplot2` does not mistake the numeric values as a continuous scale:


```{r}
names(coltab) <- 0:255 # 1:256 would be wrong by 1

ggplot() +
  geom_raster(data = map_1b_df, aes(x,y, fill = factor(values))) +
  scale_fill_manual(values = coltab) +
  theme(legend.position = "none") +
  coord_equal()

```

Note that there are thee ways to plot this type of data in `ggplot2`: `geom_rect`, `geom_tile` and `geom_raster`. The latter is "a high performance special case for when all the tiles are the same size" (see [the docs](https://ggplot2.tidyverse.org/reference/geom_tile.html).


Alternatively, you can add the hex colourcode to the dataframe and use `scale_fill_identitiy()`.
```{r, eval = FALSE}
map_1b_df <- mutate(map_1b_df, colorvalue = coltab[values+1])
ggplot() +
  geom_raster(data = map_1b_df, aes(x,y, fill = colorvalue)) +
  scale_fill_identity() +
  theme(legend.position = "none") +
  coord_equal()
```

The big drawback of using `ggplot2` in this way: you cannot add a second layer with a `fill` aesthetic, since `ggplot2` only allows one per plot.

###### Multiple bands



In case of a 3-Band Raster Brick, each layer holds a value of 0 to 255 representing colours in the RGB colour model. Usually, the first layer represents **R**ed, the second **G**reen and the third **B**lue. We can subset the `RasterBrick` into the individual `RasterLayers` using double brackets (`[[1]]` for the first layer).

```{r}
map_3b

map_3b[[1]]
```

As with the singleband raster, we can extract the Values (this time RGB-Values) of each layer with `getValues()`.

```{r}
head(values(map_3b[[1]]),50)
```

Again like with the single band raster, we now transform the Raster into a dataframe, this time storing the rgb values in columns (instead of the hex values as before).

```{r}
map_3b_df <- data.frame(
  coordinates(map_3b),
  r = values(map_3b[[1]]), 
  g = values(map_3b[[2]]),
  b = values(map_3b[[3]]))

head(map_3b_df)
```


With the function `rbg()` we can transform values into hex-codes.
```{r}
rgb(10,10,10,maxColorValue = 255)

ggplot() + 
  geom_raster(data = map_3b_df, aes(x,y,fill = rgb(r,g,b,maxColorValue = 255))) + 
  scale_fill_identity() +
  coord_equal()
```

Tipp: if your file is to large and ggplot has trouble plotting it fast, you can use `aggregate` to reduce the resolution

```{r}
map_3b_lowres <- raster::aggregate(map_3b, fact = 5)
```



##### Greyscale

In greyscale maps, `raster` cannot find a associated colortable to match the values to rgb or hex colours.
```{r}
map_grey_path <- "sample_data/SMR_Musterdaten/SMR50_LV95_KGRS_Mosaic.tif"
map_grey <- raster(map_grey_path)

head(getValues(map_grey))
head(colortable(map_grey))
```

However, the package `tiff` _can_ find the rgb values of a tiff, we just need to set `indexed = TRUE` and get the `color.map` attribute of the output:

```{r}
library(tiff)

index <- readTIFF(map_grey_path,indexed = TRUE)
colormap <- attr(index,"color.map")


colormap[,1:8] # only showing the first 8 columns (of 256)
```

This is an index with three rows (**R**ed, **G**reen and **B**lue) and 265 columns (with are the Values from `getValues()`). We can now use `rgb()` to turn this matrix into a vector of 265 hex colours:

```{r}
coltab <- rgb(colormap[1,],colormap[2,],colormap[3,])

scales::show_col(coltab,labels = FALSE)
```

Like before, we now turn the `RasterLayer` into a dataframe and use the `Values` and the newly created colortable to assign the correct values:

```{r}
map_grey_df <- data.frame(coordinates(map_grey),values = getValues(map_grey))


names(coltab) <- 0:255

ggplot() +
  geom_raster(data = map_grey_df, aes(x,y, fill = factor(values))) +
  scale_fill_manual(values = coltab) +
  theme(legend.position = "none") +
  coord_equal()
```


